---
title: "Model Replication"
author: "Kyle Rohde, Caroline Devlin"
date: "9/10/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this document is to replicate "The Impact of Covid-19 Testing on College Campuses" research done by Zhuoting Yu, Akane B. Fujimoto, Pinar Keskinocak, and Julie Swann.

The model created considers various states to describe a student's status as it relates to Covid-19 infection. The states are as follows:

$S_v$: susceptible and vaccinated

$S_n$: susceptible and not vaccinated

$I_{a,v}$: infected, asymptomatic, and vaccinated

$I_{a,n}$: infected, asymptomatic, and not vaccinated

$I_{s,v}$: infected, symptomatic, and vaccinated

$I_{s,n}$: infected, symptomatic, and not vaccinated

$TFN_{a(s)}$: infected and asymptomatic (symptomatic), has been tested with false negative results

$TISO^+_{a(s)}$: infected and asymptomatic (symptomatic), has been tested and currently in isolation

$TISO^-_{a(s)}$: infected and asymptomatic (symptomatic), has been tested and currently not in isolation

$RC$: recovered and confirmed, i.e., has been diagnosed previously

$RU$: recovered and unknown, i.e., has not been diagnosed previously

The following equations describe the fraction of individuals in each state over time **while considering varying masking compliance situations**:

$$
\dot{S_v} = -\beta_1(1-MC)S_v(I_{a,v} + TFN_a + TISO_a^-)-\beta_2(1-MC)S_v(I_{s,v}+TFN_s+TISO_s^-)
$$

$$
\dot{S_n} = -\beta_3(1-MC)S_n(I_{a,n} + TFN_a + TISO_a^-)-\beta_4(1-MC)S_n(I_{s,n}+TFN_s+TISO_s^-)
$$

$$
\dot{I_{a, v}}= (1-p_0) \times [\beta_1(1-MC)S_v(I_{a,v}+TFN_a+TISO_a^-)+\beta_2(1-MC)S_v(I_{s,v}+TFN_s+TISO_s^-)]-T_{pcr} \times I_{a,v}-T{rap} \times I_{a,v} -\gamma_aI_{a,v}
$$

$$
\dot{I_{a, n}}= (1-p_0) \times [\beta_3(1-MC)S_n(I_{a,n}+TFN_a+TISO_a^-)+\beta_4(1-MC)S_n(I_{s,n}+TFN_s+TISO_s^-)]-T_{pcr} \times I_{a,n}-T{rap} \times I_{a,n} -\gamma_aI_{a,n}
$$


$$
\dot{I_{s, v}} = p_0 \times [\beta_1(1-MC)S_v(I_{a,v}+TFN_a+TISO_a^-)+\beta_2(1-MC)S_v(I_{s,v}+TFN_s+TISO_s^-)] − T_{pcr} × I_{s,v} − T_{rap} × I_{s,v} − γ_sI_{s,v}
$$
$$
\dot{I_{s, v}} = p_0 \times [\beta_3(1-MC)S_n(I_{a,n}+TFN_a+TISO_a^-)+\beta_4(1-MC)S_n(I_{s,n}+TFN_s+TISO_s^-)] − T_{pcr} × I_{s,n} − T_{rap} × I_{s,n} − γ_sI_{s,n}
$$


$$
\dot{TFN_a} = [T_{pcr} × FN_{pcr_a} × (I_{a,v} + I_{a,n})] + [T_{rap} × FN_{rap_a} × (I_{a,v} + I_{a,n})] − γ_a × TFN_a
$$

$$
\dot{TFN_s} = [T_{pcr} × FN_{pcr_s} × (I_{s,v} + I_{s,n})]+ [T_{rap} × FN_{rap_s} × (I_{s,v} + I_{s,n})] − γ_s × TFN_s
$$

$$
\dot{TISO^+_a} =C_a × [T_{pcr} × FN_{pcr_a} × (I_{a,v} + I_{a,n}) + T_{rap} × FN_{rap_a} × (I_{a,v} + I_{a,n})] − γ_a × TISO^+_a
$$

$$
\dot{TISO^+_s} =C_s × [T_{pcr} × FN_{pcr_s} × (I_{s,v} + I_{s,n}) + T_{rap} × FN_{rap_s} × (I_{s,v} + I_{s,n})] − γ_s × TISO^+_s
$$

$$
\dot{TISO^−_a} =(1 − C_a) × [T_{pcr} × FN_{pcr_a} × (I_{a,v} + I_{a,n}) + T_{rap} × FN_{rap_a} × (I_{a,v} + I_{a,n})] − γ_a × TISO^−_a
$$

$$
\dot{TISO^−_s} =(1 − C_s) × [T_{pcr} × FN_{pcr_s} × (I_{s,v} + I_{s,n}) + T_{rap} × FN_{rap_s} × (I_{s,v} + I_{s,n})] − γ_s × TISO^−_s
$$
$$
\dot{RC} = \gamma × [TISO_a^+ + TISO_s^+ + TISO_a^- + TISO_s^-]
$$

$$
\dot{RU} = \gamma × [(I_{a,v} + I_{a,n}) + (I_{s,v} + I_{s,n}) + TFN_a + TFN_s]
$$


Below are the variable definitions

|          Parameter         |                                             Description                                             |           Value          |
|:--------------------------:|:---------------------------------------------------------------------------------------------------:|:------------------------:|
|         $\beta_1$          | Transmission rate due to contact between susceptible, vaccinated and asymptomatic infected          |            0.1           |
|         $\beta_2$          | Transmission rate due to contact between susceptible, vaccinated and symptomatic infected           |            0.1           |
|         $\beta_3$          | Transmission rate due to contact between susceptible, not vaccinated and asymptomatic infected      |            0.1           |
|         $\beta_4$          | Transmission rate due to contact between susceptible, not vaccinated and symptomatic infected       |            0.1           |
|          $\gamma$          |                              Rate of recovery of an infected individual                             |            0.1           |
|            $p_0$           |                            Proportion of infections that are symptomatic                            |            0.5           |
|         $C_a$,$C_s$        |        Isolation compliance rate for asymptomatic/symptomatic infected and tested individuals       |         0.7, 0.9         |
|  $FN_{pcr_a}$,$FN_{pcr_s}$ |               False negative rate of a PCR test for asymptomatic/symptomatic patients               |         15%, 15%         |
| $FN_{rap_a}$, $FN_{rap_s}$ |              False negative rate of a rapid test for asymptomatic/symptomatic patients              |         30%, 30%         |
| $FP_{pcr_a}$, $FP_{pcr_s}$ |               False positive rate of a PCR test for asymptomatic/symptomatic patients               |          5%, 5%          |
| $FP_{rap_a}$, $FP_{rap_s}$ |               False positive ate of a rapid test for asymptomatic/symptomatic patients              |         10%, 10%         |
|          $T_{pcr}$         |                         Percentage of students that have access to PCR tests                        | Defined in each scenario |
|  $T_{rap_a}$, $T_{rap_s}$  |                    Percentage of students that have access to rapid antigen tests                   | Defined in each scenario |
|          $MC$              |                    Level of mask efficacy dependent on school's masking policy                      | Defined in each scenario |
|          $VR_s$            |                    Vaccination Rate of students in school defined by school administrator           | Defined in each scenario |
|          $VR_f$            |                    Vaccination Rate of students in school defined by school administrator           | Defined in each scenario |


Below is the R code for defining the initial values and constants of the system.
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
# Dependencies
library(hash)
library(deSolve)
library(scatterplot3d)
library(lubridate)
library(ggplot2)
```

```{r}
# Download data from GDPH website
temp_file <- tempfile()
download.file("https://ga-covid19.ondemand.sas.com/docs/ga_covid_data.zip", temp_file)
data <- read.csv(unz(temp_file, "county_cases.csv"))
unlink(temp_file)

# Remove unused columns
county_data <- subset(data, select = -c(county_id, 
                                 State.FIPS.code,
                                 County.FIPS.code,
                                 confirmed_case_hospitalization,
                                 antigen_case_hospitalization,
                                 deaths,
                                 case.rate,
                                 death.rate,
                                 X14.day.case.rate,
                                 antigen_cases,
                                 probable_deaths))
county_data <- county_data[-c(108, 146),]

# Account for 1 in 4 cases being reported to DPH
county_data$X14.day.cases <- ceiling(county_data$X14.day.cases * runif(1, min = 3.4, max = 4.7))

attach(county_data)
```

```{r}
# Hard coded all counties in Georgia
counties <- hash(
                  keys=c(
                   "Appling", "Atkinson", "Bacon", "Baker", "Baldwin", "Banks", "Barrow", "Bartow",
                   "Ben Hill", "Berrien", "Bibb", "Bleckley", "Brantley", "Brooks", "Bryan", "Bulloch",
                   "Burke", "Butts", "Calhoun", "Camden", "Candler", "Carroll", "Catoosa", "Charlton",
                   "Chatham", "Chattahoochee", "Chattooga", "Cherokee", "Clarke", "Clay", "Clayton",
                   "Clinch", "Cobb", "Coffee", "Colquitt", "Columbia", "Cook", "Coweta", "Crawford", "Crisp",
                   "Dade", "Dawson", "DeKalb", "Decatur", "Dodge", "Dooly", "Dougherty", "Douglas", "Early",
                   "Echols", "Effingham", "Elbert", "Emanuel", "Evans", "Fannin", "Fayette", "Floyd",
                   "Forsyth", "Franklin", "Fulton", "Gilmer", "Glascock", "Glynn", "Gordon", "Grady",
                   "Greene", "Gwinnett", "Habersham", "Hall", "Hancock", "Haralson", "Harris", "Hart",
                   "Heard", "Henry", "Houston", "Irwin", "Jackson", "Jasper", "Jeff Davis", "Jefferson",
                   "Jenkins", "Johnson", "Jones", "Lamar", "Lanier", "Laurens", "Lee", "Liberty", "Lincoln",
                   "Long", "Lowndes", "Lumpkin", "Macon", "Madison", "Marion", "McDuffie", "McIntosh", 
                   "Meriwether", "Miller", "Mitchell", "Monroe", "Montgomery", "Morgan", "Murray", 
                   "Muscogee", "Newton", "Oconee", "Oglethorpe", "Paulding", "Peach", "Pickens", "Pierce", 
                   "Pike", "Polk", "Pulaski", "Putnam", "Quitman", "Rabun", "Randolph", "Richmond", 
                   "Rockdale", "Schley", "Screven", "Seminole", "Spalding", "Stephens", "Stewart", "Sumter",
                   "Talbot", "Taliaferro", "Tattnall", "Taylor", "Telfair", "Terrell", "Thomas", "Tift", 
                   "Toombs", "Towns", "Treutlen", "Troup", "Turner", "Twiggs", "Union", "Upson", "Walker",
                   "Walton", "Ware", "Warren", "Washington", "Wayne", "Webster", "Wheeler", "White", 
                   "Whitfield", "Wilcox", "Wilkes", "Wilkinson", "Worth"
                  ),
                  values=1:159
                )

generate_school <- function(county, days, masking, student_pop, faculty_pop, student_vax, faculty_vax, testing, pcr, elem) {
  school <- hash()
  school["county"]           <- county
  school["days"]             <- days
  school["masking"]          <- masking * (1/6)
  school["student_pop"]      <- student_pop
  school["faculty_pop"]      <- faculty_pop
  school["student_vax"]      <- student_vax
  school["faculty_vax"]      <- faculty_vax
  school["testing"]          <- testing
  school["pcr"]              <- pcr
  school["elem"]             <- elem
  return(school)
}
```


```{r}
# Functions defining initial states for all counties
p0 = 0.7  # proportion of infections that are symptomatic

initial_susceptibility <- function(county_data) {
  result <- (county_data$population - county_data$cases) / county_data$population
  return(result)
}

initial_infected_symptomatic <- function(county_data) {
  result <- ((county_data$X14.day.cases) / county_data$population) * p0
  return(result)
}

initial_infected_asymptomatic <- function(county_data) {
  result <- ((county_data$X14.day.cases) / county_data$population) * (1 - p0)
  return(result)
}

initial_recovered_unknown <- function(county_data) {
  result <- (county_data$cases - county_data$X14.day.cases) / county_data$population
  return(result)
}

county <- "Fulton"
days <- 14
masking <- 0.5
testing <- 0.5
pcr <- TRUE
student_pop <- 1400
faculty_pop <- 14
VR_s <- 0.1
VR_f <- 0.2
elem <- FALSE

students_vaccinated <- student_pop * VR_s
faculty_vaccinated <- faculty_pop * VR_f
total_school_population <- student_pop + faculty_pop
percent_school_vaccinated <- (students_vaccinated + faculty_vaccinated) / total_school_population

# Scale county_data for all counties to include initial states for simulation
county_data$S_0 <- initial_susceptibility(county_data)
county_data$S_0_v <- (county_data$S_0)*percent_school_vaccinated
county_data$S_0_n <- (county_data$S_0)*(1 - percent_school_vaccinated)
county_data$I_sv <- (percent_school_vaccinated) * initial_infected_symptomatic(county_data)
county_data$I_sn <- (1 - percent_school_vaccinated) * initial_infected_symptomatic(county_data)
county_data$I_av <- (percent_school_vaccinated) * initial_infected_asymptomatic(county_data)
county_data$I_an <- (1 - percent_school_vaccinated) * initial_infected_asymptomatic(county_data)
county_data$R_u <- initial_recovered_unknown(county_data)

school <- generate_school(toString(county), days, masking, student_pop, faculty_pop, VR_s, VR_f, testing, pcr, elem)
single_county_data <- county_data[values(counties, key=c(county)),]
```

```{r}
# Initial built-in parameters
# The 40 found in the beta initialization signifies the simulation will run 40 times. If you would like to run it for a different number of iterations, change the values here.
b_1 = runif(40, 0.051, 0.068)          # transmission rate due to contact between susceptible vaccinated and infected asymptomatic patient
b_2 = runif(40, 0.102, 0.136)          # transmission rate due to contact between susceptible vaccinated and infected symptomatic patient
b_3 = runif(40, 0.150, 0.200)          # transmission rate due to contact between susceptible not vaccinated and infected asymptomatic patient
b_4 = runif(40, 0.300, 0.400)          # transmission rate due to contact between susceptible not vaccinated and infected symptomatic patient
gamma = 0.1        # rate of recovery for an infected individual
p0 = 0.7           # proportion of infections that are symptomatic
c_a = 1            # isolation compliance asymptomatic infected + tested patients
c_s = 1            # isolation compliance symptomatic infected + tested patients
FNpcr_a = 0.149    # false negative rate of a PCR test for asymptomatic patients
FNpcr_s = 0.149    # false negative rate of a PCR test for symptomatic patients
FNrap_a = 0.37     # false negative rate of a rapid test for asymptomatic patients
FNrap_s = 0.37     # false negative rate of a rapid test for symptomatic patients
FPpcr_a = 0.05     # false positive rate of a PCR test for asymptomatic patients
FPpcr_s = 0.05     # false positive rate of a PCR test for symptomatic patients
FPrap_a = 0.10     # false positive rate of a rapid test for asymptomatic patients
FPrap_s = 0.10     # false positive rate of a rapid test for symptomatic patients

# Initial susceptibility values
S_0_v = as.numeric(single_county_data[6])    # initial susceptibility of the vaccinated school population
S_0_n = as.numeric(single_county_data[7])    # initial susceptibility of the not vaccinated school population

# Calculations for the initial infected
I_sv = as.numeric(single_county_data[8])     # initial infected, symptomatic, and vaccinated
I_sn = as.numeric(single_county_data[9])     # initial infected, symptomatic, and not vaccinated
I_av = as.numeric(single_county_data[10])    # initial infected, asymptomatic, and vaccinated
I_an = as.numeric(single_county_data[11])    # initial infected, asymptomatic, and not vaccinated

# Recovered states
R_u = as.numeric(single_county_data[12])     # initial recovered unknown
R_c = 0.0                                    # initial recovered confirmed

# Initial test values (no tests administered)
TFN_a = 0       # initial population with a false negative and asymptomatic
TFN_s = 0       # initial population with a false negative and symptomatic 
TISO_an = 0     # initial population tested with infection, not in isolation, asymptomatic 
TISO_ap = 0     # initial population tested with infection, in isolation, asymptomatic 
TISO_sn = 0     # initial population tested with infection, not in isolation, symptomatic
TISO_sp = 0     # initial population tested with infection, in isolation, symptomatic

n = values(school, key=c("student_pop")) + values(school, key=c("faculty_pop"))       # school population

# Days in simulation
days <- values(school, key=c("days"))

# Sequence for semester
semester <- seq(1, days, by=1)

# Here we determine if we are pcr testing based on user input, if we are, assign pcr testing percentage to user inputted testing percentage. else, assign to rapid testing percentage
if (values(school, key=c("pcr"))) {
  T_pcr <- values(school, key=c("testing"))
  T_rap <- 0
} else {
  T_pcr <- 0
  T_rap <- values(school, key=c("testing"))
}

# Here we determine if the school is an elementary school based on user input. If it is, assign the elementary school multiplier to 0.73 to account for historically less transmission in elementary schools
if (values(school, key=c("elem"))) {
  elementary_multiplier <- 0.73
} else {
  elementary_multiplier <- 1
}

# Initial built in parameters for testing and no testing days
testing <- c(b_1 = b_1[1], b_2 = b_2[1], b_3 = b_3[1], b_4 = b_4[1], p0 = p0, c_a = c_a, c_s = c_s, MC = values(school, key=c("masking")), T_pcr = T_pcr, T_rap = T_rap, gamma = gamma, FNpcr_a = FNpcr_a, FNpcr_s = FNpcr_s, FNrap_a = FNrap_a, FNrap_s = FNrap_s)

no_testing <- c(b_1 = b_1[1], b_2 = b_2[1], b_3 = b_3[1], b_4 = b_4[1], p0 = p0, c_a = c_a, c_s = c_s, MC = values(school, key=c("masking")), T_pcr = 0, T_rap = 0, gamma = gamma, FNpcr_a = FNpcr_a, FNpcr_s = FNpcr_s, FNrap_a = FNrap_a, FNrap_s = FNrap_s)

# Initial state of the system
initial_state <- c(S_0_v = S_0_v, S_0_n = S_0_n, I_av = I_av, I_an = I_an, I_sv = I_sv, I_sn = I_sn, TFN_a = TFN_a, TFN_s = TFN_s, TISO_an = TISO_an, TISO_sn = TISO_sn, TISO_ap = TISO_ap, TISO_sp = TISO_sp, RC = R_c, RU = R_u)

# Initial built in parameters for testing and no testing days (no intervention strategies)
testing_baseline <- c(b_1 = b_1[1], b_2 = b_2[1], b_3 = b_3[1], b_4 = b_4[1], p0 = p0, c_a = c_a, c_s = c_s, MC.masking = 0, T_pcr = 0, T_rap = 0, gamma = gamma, FNpcr_a = FNpcr_a, FNpcr_s = FNpcr_s, FNrap_a = FNrap_a, FNrap_s = FNrap_s)

no_testing_baseline <- c(b_1 = b_1[1], b_2 = b_2[1], b_3 = b_3[1], b_4 = b_4[1], p0 = p0, c_a = c_a, c_s = c_s, MC.masking = 0, T_pcr = 0, T_rap = 0, gamma = gamma, FNpcr_a = FNpcr_a, FNpcr_s = FNpcr_s, FNrap_a = FNrap_a, FNrap_s = FNrap_s)

# Initial state of the system (no intervention strategies)
initial_state_baseline <- c(S_0_v = 0, S_0_n = S_0_n + S_0_v, I_av = 0, I_an = I_an + I_av, I_sv = 0, I_sn = I_sn + I_sv, TFN_a = TFN_a, TFN_s = TFN_s, TISO_an = TISO_an, TISO_sn = TISO_sn, TISO_ap = TISO_ap, TISO_sp = TISO_sp, RC = R_c, RU = R_u)
```


```{r}
# State transition differential equation definitions

S_dot_v <- function(b_1, MC, S_0_v, I_av, TFN_a, TISO_an, b_2, I_sv, TFN_s, TISO_sn) {
  result <- (-b_1 * (1 - MC) * S_0_v * (I_av + TFN_a + TISO_an)) - (b_2 * (1 - MC) * S_0_v * (I_sv + TFN_s + TISO_sn))
  return(result)
}

S_dot_n <- function(b_3, MC, S_0_n, I_an, TFN_a, TISO_an, b_4, I_sn, TFN_s, TISO_sn) {
  result <- (-b_3 * (1 - MC) * S_0_n * (I_an + TFN_a + TISO_an)) - (b_4 * (1 - MC) * S_0_n * (I_sn + TFN_s + TISO_sn))
  return(result)
}

I_av_dot <- function(p0, b_1, MC, S_0_v, I_av, TFN_a, TISO_an, b_2, I_sv, TFN_s, TISO_sn, T_pcr, T_rap, gamma) {
  result <- (1 - p0) * ((b_1 * (1 - MC) * S_0_v * (I_av + TFN_a + TISO_an)) + (b_2 * (1 - MC) * S_0_v * (I_sv + TFN_s + TISO_sn))) - (T_pcr * I_av) - (T_rap * I_av) - (gamma * I_av)
  return(result)
}

I_an_dot <- function(p0, b_3, MC, S_0_n, I_an, TFN_a, TISO_an, b_4, I_sn, TFN_s, TISO_sn, T_pcr, T_rap, gamma) {
  result <- (1 - p0) * ((b_3 * (1 - MC) * S_0_n * (I_an + TFN_a + TISO_an)) + (b_4 * (1 - MC) * S_0_n * (I_sn + TFN_s + TISO_sn))) - (T_pcr * I_an) - (T_rap * I_an) - (gamma * I_an)
  return(result)
}

I_sv_dot <- function(p0, b_1, MC, S_0_v, I_av, TFN_a, TISO_an, b_2, I_sv, TFN_s, TISO_sn, T_pcr, T_rap, gamma) {
  result <- p0 * ((b_1 * (1 - MC) * S_0_v * (I_av + TFN_a + TISO_an)) + (b_2 * (1 - MC) * S_0_v * (I_sv + TFN_s + TISO_sn))) - (T_pcr * I_sv) - (T_rap * I_sv) - (gamma * I_sv)
  return(result)
}

I_sn_dot <- function(p0, b_3, MC, S_0_n, I_an, TFN_a, TISO_an, b_4, I_sn, TFN_s, TISO_sn, T_pcr, T_rap, gamma) {
  result <- p0 * ((b_3 * (1 - MC) * S_0_n * (I_an + TFN_a + TISO_an)) + (b_4 * (1 - MC) * S_0_n * (I_sn + TFN_s + TISO_sn))) - (T_pcr * I_sn) - (T_rap * I_sn) - (gamma * I_sn)
  return(result)
}

TFN_a_dot <- function(T_pcr, FNpcr_a, I_av, I_an, T_rap, FNrap_a, gamma, TFN_a) {
  result <- (T_pcr * FNpcr_a * (I_av + I_an)) + (T_rap * FNrap_a * (I_av + I_an)) - (gamma * TFN_a)
  return(result)
}

TFN_s_dot <- function(T_pcr, FNpcr_s, I_sv, I_sn, T_rap, FNrap_s, gamma, TFN_s) {
  result <- (T_pcr * FNpcr_s * (I_sv + I_sn)) + (T_rap * FNrap_s * (I_sv + I_sn)) - (gamma * TFN_s)
  return(result)
}

TISO_ap_dot <- function(c_a, T_pcr, FNpcr_a, I_av, I_an, T_rap, FNrap_a, gamma, TISO_ap) {
  result <- c_a * ((T_pcr * (1 -FNpcr_a) * (I_av + I_an)) + (T_rap * (1 - FNrap_a) * (I_av + I_an))) - (gamma * TISO_ap)
  return(result)
}

TISO_sp_dot <- function(c_s, T_pcr, FNpcr_s, I_sv, I_sn, T_rap, FNrap_s, gamma, TISO_sp) {
  result <- c_s * ((T_pcr * (1 - FNpcr_s) * (I_sv + I_sn)) + (T_rap * (1 - FNrap_s) * (I_sv + I_sn))) - (gamma * TISO_sp)
  return(result)
}

TISO_an_dot <- function(c_a, T_pcr, FNpcr_a, I_av, I_an, T_rap, FNrap_a, gamma, TISO_an) {
  result <- (1 - c_a) * ((T_pcr * (1 - FNpcr_a) * (I_av + I_an)) + (T_rap * (1 - FNrap_a) * (I_av + I_an))) - (gamma * TISO_an)
  return(result)
}

TISO_sn_dot <- function(c_s, T_pcr, FNpcr_s, I_sv, I_sn, T_rap, FNrap_s, gamma, TISO_sn) {
  result <- (1 - c_s) * ((T_pcr * (1 - FNpcr_s) * (I_sv + I_sn)) + (T_rap * (1 - FNrap_s) * (I_sv + I_sn))) - (gamma * TISO_sn)
  return(result)
}

RC_dot <- function(gamma, TISO_ap, TISO_sp, TISO_an, TISO_sn) {
  result <- gamma * (TISO_ap + TISO_sp + TISO_an + TISO_sn)
  return(result)
}

RU_dot <- function(gamma, I_av, I_an, I_sv, I_sn, TFN_a, TFN_s) {
  result <- gamma * ((I_av + I_an) + (I_sv + I_sn) + TFN_a + TFN_s)
  return(result)
}
```

```{r}
# Define generic differential update function
differentials <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    dS_v <- S_dot_v(b_1, MC.masking, S_0_v, I_av, TFN_a, TISO_an, b_2, I_sv, TFN_s, TISO_sn)
    dS_n <- S_dot_n(b_3, MC.masking, S_0_n, I_an, TFN_a, TISO_an, b_4, I_sn, TFN_s, TISO_sn)
    dI_av <- I_av_dot(p0, b_1, MC.masking, S_0_v, I_av, TFN_a, TISO_an, b_2, I_sv, TFN_s, TISO_sn, T_pcr, T_rap, gamma)
    dI_an <- I_an_dot(p0, b_3, MC.masking, S_0_n, I_an, TFN_a, TISO_an, b_4, I_sn, TFN_s, TISO_sn, T_pcr, T_rap, gamma)
    dI_sv <- I_sv_dot(p0, b_1, MC.masking, S_0_v, I_av, TFN_a, TISO_an, b_2, I_sv, TFN_s, TISO_sn, T_pcr, T_rap, gamma)
    dI_sn <- I_sn_dot(p0, b_3, MC.masking, S_0_n, I_an, TFN_a, TISO_an, b_4, I_sn, TFN_s, TISO_sn, T_pcr, T_rap, gamma)
    dTFN_a <- TFN_a_dot(T_pcr, FNpcr_a, I_av, I_an, T_rap, FNrap_a, gamma, TFN_a)
    dTFN_s <- TFN_s_dot(T_pcr, FNpcr_s, I_sv, I_sn, T_rap, FNrap_s, gamma, TFN_s)
    dTISO_ap <- TISO_ap_dot(c_a, T_pcr, FNpcr_a, I_av, I_an, T_rap, FNrap_a, gamma, TISO_ap)
    dTISO_sp <- TISO_sp_dot(c_s, T_pcr, FNpcr_s, I_sv, I_sn, T_rap, FNrap_s, gamma, TISO_sp)
    dTISO_an <- TISO_an_dot(c_a, T_pcr, FNpcr_a, I_av, I_an, T_rap, FNrap_a, gamma, TISO_an)
    dTISO_sn <- TISO_sn_dot(c_s, T_pcr, FNpcr_s, I_sv, I_sn, T_rap, FNrap_s, gamma, TISO_sn)
    dRC <- RC_dot(gamma, TISO_ap, TISO_sp, TISO_an, TISO_sn)
    dRU <- RU_dot(gamma, I_av, I_an, I_sv, I_sn, TFN_a, TFN_s)
    return(list(c(dS_v, dS_n, dI_av, dI_an, dI_sv, dI_sn, dTFN_a, dTFN_s, dTISO_an, dTISO_sn, dTISO_ap, dTISO_sp, dRC, dRU)))
  })
}

run_simulation <- function(state, testing, no_testing, semester) {
  states <- list()
  for (t in semester) {
    day <- seq(t - 1, t, by=1)
    if (t %% 7 == 0) {
      parameters <- testing
    } else {
      parameters <- no_testing
    }
    state <- ode(y = state, times = day, func = differentials, parms = parameters)[2,c(2:15)]
    states[[length(states) + 1]] <- state
  }
  return(states)
}

# Helper function for internal use
update_parms <- function(i, testing, no_testing) {
  testing["b_1"] <- b_1[i]
  testing["b_2"] <- b_2[i]
  testing["b_3"] <- b_3[i]
  testing["b_4"] <- b_4[i]
  no_testing["b_1"] <- b_1[i]
  no_testing["b_2"] <- b_2[i]
  no_testing["b_3"] <- b_3[i]
  no_testing["b_4"] <- b_4[i]
  return(c(testing, no_testing))
}

# Use this function to run simulation for j number of times
run_simulation_repeated <- function(j, state, testing, no_testing, semester) {
  all_states <- list()
  for (i in 1:j) {
    parms <- update_parms(i, testing, no_testing)
    solution <- run_simulation(state, parms[1:15], parms[16:31], semester)
    all_states[[length(all_states) + 1]] <- solution
  }
  return(all_states)
}
```

The following code block is primarily used for statistical analysis of the results from running the simulation $i$ number of times. By default, $i = 40$. To change the total number of iterations, please modify the $40$ found in the function call on the last line of the previous code block.
```{r}
# Function to determine number of cases for each day. Returns list of length days where each element is the cases for that day
calculate_cases <- function(states) {
  cases_per_day <- list()
  cases_per_day[1] <- 0
  for (i in 2:days) {
    previous_state <- states[[i - 1]]
    current_state <- states[[i]]
    
    current_susceptibility <- as.numeric(current_state["S_0_v"] + current_state["S_0_n"])
    infection_attack_rate <- as.numeric((previous_state["S_0_v"] + previous_state["S_0_n"])) - current_susceptibility
    decimal_of_IAR <- as.numeric((infection_attack_rate * n) - floor(infection_attack_rate * n))
    if (decimal_of_IAR >= 0.7) {
      cases_per_day[i] <- ceiling(infection_attack_rate * n) * elementary_multiplier
    } else {
      cases_per_day[i] <- floor(infection_attack_rate * n) * elementary_multiplier
    }
    cases_per_day[i] <- as.numeric(cases_per_day[i])
  }
  return(cases_per_day)
}

# Function to determine cumulative cases over the simulation duration. Returns list of length days where each element is the total cases since the start of the simulation
calculate_cumulative_cases <- function(cases_per_day) {
  cumulative_cases <- list()
  running_total <- 0
  for (i in 1:length(cases_per_day)) {
    cumulative_cases[i] <- running_total
    running_total <- running_total + as.numeric(cases_per_day[i])
  }
  return(cumulative_cases)
}

# Use this function to get cases per day for multiple simulation runs
calculate_cases_multiple_simulations <- function(j, results) {
  total_cases <- list()
  for (i in 1:j) {
    cases_per_day <- calculate_cases(results[[i]])
    total_cases[[length(total_cases) + 1]] <- cases_per_day
  }
  return(total_cases)
}

# Function converts total_cases to a list of list. Outer list is number of cases for day i within each simulation. Inner list is of length number of simulations
reorganize_total_cases <- function(n, total_cases) {
  reorganized_cases <- list()
  for (j in 1:days) {
    cases_day_j <- list()
    for (i in 1:n) {
      cases_day_j[[length(cases_day_j) + 1]] <- total_cases[[i]][[j]]
    }
    reorganized_cases[[length(reorganized_cases) + 1]] <- cases_day_j
  }
  return(reorganized_cases)
}


# Use this function to convert reorganized cases to cumulative cases over i simulations
reorganize_cumulative_cases <- function(n, total_cases) {
  temp_cumulative_cases <- list()
  for (i in 1:n) {
    temp_cumulative_cases[[length(temp_cumulative_cases) + 1]] <- calculate_cumulative_cases(total_cases[[i]])
  }
  
  cumulative_cases <- list()
  for (j in 1:days) {
    cumulative_cases_day_j <- list()
    for (i in 1:n) {
      cumulative_cases_day_j[[length(cumulative_cases_day_j) + 1]] <- temp_cumulative_cases[[i]][[j]]
    }
    cumulative_cases[[length(cumulative_cases) + 1]] <- cumulative_cases_day_j
  }
  return(cumulative_cases)
}

# Get average for cases per day
avg_cases_per_day <- function(reorganized_cases) {
  day_avgs <- list()
  for (i in 1:days) {
    avg_case <- mean(as.numeric(reorganized_cases[[i]]))
    avg_case_decimal <- avg_case - floor(avg_case)
    if (avg_case_decimal >= 0.5) {
      avg_case_int <- as.numeric(ceiling(avg_case))
    } else {
      avg_case_int <- as.numeric(floor(avg_case))
    }
    day_avgs[[length(day_avgs) + 1]] <- avg_case_int
  }
  return(day_avgs)
}

# Get standard deviation for cases per day
st_dev_cases_per_day <- function(reorganized_cases) {
  st_devs <- list()
  for (i in 1:days) {
    st_devs[[length(st_devs) + 1]] <- sd(as.numeric(reorganized_cases[[i]]))
  }
  return(st_devs)
}

# Get margins for confidence intervals for cases per day
margins_per_day <- function(reorganized_cases) {
  n <- length(reorganized_cases[[1]])
  margins <- list()
  for (i in 1:days) {
    st_dev <- sd(as.numeric(reorganized_cases[[i]]))
    margins[[length(margins) + 1]] <- qt(0.975, df=n-1) * st_dev / sqrt(n)
  }
  return(margins)
}

# Use this function to graph average cases per day over days in simulation. Returns graph as object
generate_graph_cases_per_day <- function(reorganized_cases, reorganized_cases_baseline) {
  n <- length(reorganized_cases[1])
  avgs <- avg_cases_per_day(reorganized_cases)
  st_devs <- st_dev_cases_per_day(reorganized_cases)
  margins <- margins_per_day(reorganized_cases)
  upper <- as.numeric(avgs) - as.numeric(margins)
  lower <- as.numeric(avgs) + as.numeric(margins)
  
  avgs_base <- avg_cases_per_day(reorganized_cases_baseline)
  st_devs_base <- st_dev_cases_per_day(reorganized_cases_baseline)
  margins_base <- margins_per_day(reorganized_cases_baseline)
  upper_base <- as.numeric(avgs_base) - as.numeric(margins_base)
  lower_base <- as.numeric(avgs_base) + as.numeric(margins_base)
  
  x <- 1:days
  y <- as.numeric(avg_cases_per_day(reorganized_cases))
  y_base <- as.numeric(avg_cases_per_day(reorganized_cases_baseline))
  
  data <- data.frame(x, y, lower, upper)
  data_base <- data.frame(x, y_base, lower_base, upper_base)
  df <- data.frame(data, data_base)
  
  p <- ggplot(data = df, aes(x = x)) +
        geom_line(aes(y = y, colour="Intervention")) +
        geom_point(aes(x=x, y=y, colour="Intervention")) +
        geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.05) +
        geom_line(aes(y = y_base, colour="No Intervention")) +
        geom_point(aes(x=x, y=y_base, colour="No Intervention")) +
        geom_ribbon(aes(ymin = lower_base, ymax = upper_base), alpha = 0.05) +
        scale_color_manual("",
              breaks = c("Intervention", "No Intervention"),
              values = c("green", "red")) +
        scale_x_continuous(breaks=c(1:days)) +
        scale_y_continuous(breaks=c(1:(y_base[days] + 1))) +
        xlab("Days") + 
        ylab("Number of Cases") +
        labs(title="Predicted Number of new Cases per Day")
  return(p + theme_bw())
}

# Use this function to graph cumulative cases over days in simulation. Returns graph as object
generate_graph_cumulative_cases <- function(cumulative_cases, cumulative_cases_baseline) {
  n <- length(cumulative_cases[1])
  avg_cases <- as.numeric(avg_cases_per_day(cumulative_cases))
  st_devs <- as.numeric(st_dev_cases_per_day(cumulative_cases))
  margins <- as.numeric(margins_per_day(cumulative_cases))
  upper <- avg_cases + margins
  lower <- avg_cases - margins
  
  avg_cases_base <- as.numeric(avg_cases_per_day(cumulative_cases_baseline))
  st_devs_base <- as.numeric(st_dev_cases_per_day(cumulative_cases_baseline))
  margins_base <- as.numeric(margins_per_day(cumulative_cases_baseline))
  upper_base <- avg_cases_base + margins_base
  lower_base <- avg_cases_base - margins_base
  
  x <- 1:days
  y <- avg_cases
  y_base <- avg_cases_base
  
  data <- data.frame(x, y, lower, upper)
  data_base <- data.frame(x, y_base, lower_base, upper_base)
  df <- data.frame(data, data_base)
  
  y_axis <- c(0:y_base[days] + 1)
  y_axis <- which(y_axis %% 5 == 0, arr.ind = FALSE)
  
  p <- ggplot(data = df, aes(x = x)) +
        geom_line(aes(y = y, colour="Intervention")) +
        geom_point(aes(x=x, y=y, colour="Intervention")) +
        geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.05) +
        geom_line(aes(y = y_base, colour="No Intervention")) +
        geom_point(aes(x=x, y=y_base, colour="No Intervention")) +
        geom_ribbon(aes(ymin = lower_base, ymax = upper_base), alpha = 0.05) +
        scale_color_manual("",
              breaks = c("Intervention", "No Intervention"),
              values = c("green", "red")) +
        scale_fill_manual("",
              breaks = c("Intervention", "No Intervention"),
              values = c("grey41", "grey41")) +
        scale_x_continuous(breaks=c(1:days)) +
        scale_y_continuous(breaks=y_axis) +
        xlab("Days") + 
        ylab("Cumulative Cases") +
        labs(title="Predicted Number of Cumulative Cases")
  return(p + theme_bw())
}

# Run simulation for 40 repetitions
cases_per_day_multiple_simulations <- run_simulation_repeated(40, initial_state, testing, no_testing, semester)
cases_per_day_multiple_simulations_baseline <- run_simulation_repeated(40, initial_state_baseline, testing_baseline, no_testing_baseline, semester)

# Calculate case projects per day across all 40 simulations
total_cases <- calculate_cases_multiple_simulations(40, cases_per_day_multiple_simulations)
total_cases_baseline <- calculate_cases_multiple_simulations(40, cases_per_day_multiple_simulations_baseline)

# Reorganize case projection output for easier statistical analysis
reorganized_cases <- reorganize_total_cases(40, total_cases)
reorganized_cases_baseline <- reorganize_total_cases(40, total_cases_baseline)

# Reorganize case projection output for easier statistical analysis (cumulative cases)
cumulative_cases <- reorganize_cumulative_cases(40, total_cases)
cumulative_cases_baseline <- reorganize_cumulative_cases(40, total_cases_baseline)

# Generate projection graph with confidence intervals
cases_per_day_graph <- generate_graph_cases_per_day(reorganized_cases, reorganized_cases_baseline)
cumulative_cases_graph <- generate_graph_cumulative_cases(cumulative_cases, cumulative_cases_baseline)

print(cases_per_day_graph)
print(cumulative_cases_graph)
```