---
title: "Model Replication"
author: "Kyle Rohde, Caroline Devlin"
date: "9/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The purpose of this document is to replicate "The Impact of Covid-19 Testing on College Campuses" research done by Zhuoting Yu, Akane B. Fujimoto, Pinar Keskinocak, and Julie Swann.

The model created considers various states to describe a student's status as it relates to Covid-19 infection. The states are as follows:

$S$: susceptible

$I_a$: infected and asymptomatic

$I_s$: infected and symptomatic

$TFN_{a(s)}$: infected and asymptomatic (symptomatic), has been tested with false negative results

$TISO^+_{a(s)}$: infected and asymptomatic (symptomatic), has been tested and currently in isolation

$TISO^-_{a(s)}$: infected and asymptomatic (symptomatic), has been tested and currently not in isolation

$RC$: recovered and confirmed, i.e., has been diagnosed previously

$RU$: recovered and unknown, i.e., has not been diagnosed previously

The follow equations describe the fraction of individuals in each state over time

$$
S = -\beta_aS(I_a + TFN_a + TISO_a^-)-\beta_sS(I_s+TFN_s+TISO_s^-)
$$

$$
I_a= (1-p_0) \times [\beta_aS(I_a+TFN_a+TISO_a^-)+\beta_sS(I_s+TFN_s+TISO_s^-)]-T_{pcr} \times I_a-T{rap} \times I_a -\gamma_aI_a
$$

$$
I_s = p_0 \times [\beta_aS(I_a+TFN_a+TISO_a^-)+\beta_sS(I_s+TFN_s+TISO_s^-)] − T_{pcr} × I_s − T_{rap} × I_s − γ_sI_s
$$

$$
TFN_a =T_{pcr} × FN_{pcr_a} × I_a + T_{rap} × FN_{rap_a} × I_a − γ_a × TFN_a
$$

$$
TFN_s =T_{pcr} × FN_{pcr_s} × I_s + T_{rap} × FN_{rap_s} × I_s − γ_s × TFN_s
$$

$$
TISO^+_a =C_a × [T_{pcr} × FN_{pcr_a} × I_a + T_{rap} × FN_{rap_a} × I_a] − γ_a × TISO^+_a
$$

$$
TISO^+_s =C_s × [T_{pcr} × FN_{pcr_s} × I_s + T_{rap} × FN_{rap_s} × I_s] − γ_s × TISO^+_s
$$

$$
TISO^−_a =(1 − C_a) × [T_{pcr} × FN_{pcr_a} × I_a + T_{rap} × FN_{rap_a} × I_a] − γ_a × TISO^−_a
$$

$$
TISO^−_s =(1 − C_s) × [T_{pcr} × FN_{pcr_s} × I_s + T_{rap} × FN_{rap_s} × I_s] − γ_s × TISO^−_s
$$

Below are the variable definitions

|          Parameter         |                                             Description                                             |           Value          |
|:--------------------------:|:---------------------------------------------------------------------------------------------------:|:------------------------:|
|     $\beta_a$,$\beta_s$    | Transmission rate due to contact between susceptible and asymptomatic/symptomatic infected subjects |          0.1,0.2         |
|          $\gamma$          |                              Rate of recovery of an infected individual                             |            0.1           |
|            $p_0$           |                            Proportion of infections that are symptomatic                            |            0.5           |
|         $C_a$,$C_s$        |        Isolation compliance rate for asymptomatic/symptomatic infected and tested individuals       |         0.7, 0.9         |
|  $FN_{pcr_a}$,$FN_{pcr_s}$ |               False negative rate of a PCR test for asymptomatic/symptomatic patients               |         15%, 15%         |
| $FN_{rap_a}$, $FN_{rap_s}$ |              False negative rate of a rapid test for asymptomatic/symptomatic patients              |         30%, 30%         |
| $FP_{pcr_a}$, $FP_{pcr_s}$ |               False positive rate of a PCR test for asymptomatic/symptomatic patients               |          5%, 5%          |
| $FP_{rap_a}$, $FP_{rap_s}$ |               False positive ate of a rapid test for asymptomatic/symptomatic patients              |         10%, 10%         |
|          $T_{pcr}$         |                         Percentage of students that have access to PCR tests                        | Defined in each scenario |
|  $T_{rap_a}$, $T_{rap_s}$  |                    Percentage of students that have access to rapid antigen tests                   | Defined in each scenario |

* Baseline: No testing
* $PCR_{100}$: PCR tests available to 100% of eligible student body per week
* $RA_{100}$: Rapid antigen tests available to 100% of the eligible student body per week
* $PCR_{50}$: PCR tests available to 50% of the eligible student body per week
* $RA_{50}$: Rapid antigen tests available to 50% of the eligible student body per week
* $PCR_{33}$: PCR tests available to 33% of the eligible student body per week
* $RA_{33}$: Rapid antigen tests available to 33% of the eligible student body per week

```{r}
# Constants
b_a = 0.1        # transmission rate due to contact of asymptomatic patient
b_s = 0.2        # transmission rate due to contact of symptomatic patient
gamma = 0.1     # rate of recovery for an infected individual
p0 = 0.5        # proportion of infections that are symptomatic
c_a = 0.7        # isolation compliance asymptomatic infected + tested patients
c_s = 0.9        # isolation compliance symptomatic infected + tested patients
FNpcr_a = 0.15  # false negative rate of a PCR test for asymptomatic patients
FNpcr_s = 0.15  # false negative rate of a PCR test for symptomatic patients
FNrap_a = 0.30  # false negative rate of a rapid test for asymptomatic patients
FNrap_s = 0.30  # false negative rate of a rapid test for symptomatic patients
FPpcr_a = 0.05  # false positive rate of a PCR test for asymptomatic patients
FPpcr_s = 0.05  # false positive rate of a PCR test for symptomatic patients
FPrap_a = 0.10  # false positive rate of a rapid test for asymptomatic patients
FPrap_s = 0.10  # false positive rate of a rapid test for symptomatic patients

# Initial values
S_0 = 0.895     # initial susceptibility of the student body
I_s = 0.0025    # initial infected and symptomatic
I_a = 0.0025    # initial infected and asymptomatic
R_u = 0.10      # initial recovered unknown

# Initial test values (no tests administered)
TFN_a = 0       # initial population with a false negative and asymptomatic
TFN_s = 0       # initial population with a false negative and symptomatic 
TISO_an = 0     # initial population tested with infection, not in isolation, asymptomatic 
TISO_ap = 0     # initial population tested with infection, in isolation, asymptomatic 
TISO_sn = 0     # initial population tested with infection, not in isolation, symptomatic
TISO_sp = 0     # initial population tested with infection, in isolation, symptomatic

n = 15000       # student population
```

The following creates a function that can be used to calculate values for each of the 9 equations used in the "The Impact of Covid-19 Testing on College Campuses" week over week. The function allows the user to input initial values for each of the specified parameters used in the research done by Zhuoting Yu, Akane B. Fujimoto, Pinar Keskinocak, and Julie Swann, as well as input the number of weeks that the user would like to forecast for. The function generates the predicted values for the following week and then recursively calls on itself using the values predicted for the next week. This allows the model to forecast for multiple weeks at a time as opposed to manually calculating each of the values each week, as well as allows the user to input values as data becomes known (basically allowing them to reset the initial values at any point in time).

```{r}
s_dot <- function(b_a, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn) {
  result <- (-b_a * S_0 * (I_a + TFN_a + TISO_an)) - (b_s * S_0 * (I_s + TFN_s + TISO_sn))
  return(result)
}

I_a_dot <- function(p0, b_a, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn, T_pcr, T_rap, gamma) {
  result <- (1 - p0) * (b_a * S_0 * (I_a + TFN_a + TISO_an) + b_s * S_0 * (I_s + TFN_s + TISO_sn)) - T_pcr * I_a - T_rap * I_a - gamma * I_a
  return(result)
}

I_s_dot <- function(p0, b_a, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn, T_pcr, T_rap, gamma) {
  result <- p0 * (b_a * S_0 * (I_a + TFN_a + TISO_an) + b_s * S_0 * (I_s + TFN_s + TISO_sn)) - T_pcr * I_s - T_rap * I_s - gamma * I_s
  return(result)
}

TFN_a_dot <- function(T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TFN_a) {
  result <- T_pcr * FNpcr_a * I_a + T_rap * FNrap_a * I_a - gamma * TFN_a
  return(result)
}

TFN_s_dot <- function(T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TFN_s) {
  result <- T_pcr * FNpcr_s * I_s + T_rap * FNrap_s * I_s - gamma * TFN_s
  return(result)
}

TISO_ap_dot <- function(c_a, T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TISO_ap) {
  result <- c_a * (T_pcr * FNpcr_a * I_a + T_rap * FNrap_a * I_a) - gamma * TISO_ap
  return(result)
}

TISO_sp_dot <- function(c_s, T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TISO_sp) {
  result <- c_s * (T_pcr * FNpcr_s * I_s + T_rap * FNrap_s * I_s) - gamma * TISO_sp
  return(result)
}

TISO_an_dot <- function(c_a, T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TISO_an) {
  result <- (1 - c_a) * (T_pcr * FNpcr_a * I_a + T_rap * FNrap_a * I_a) - gamma * TISO_an
  return(result)
}

TISO_sn_dot <- function(c_s, T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TISO_sn) {
  result <- (1 - c_s) * (T_pcr * FNpcr_s * I_s + T_rap * FNrap_s * I_s) - gamma * TISO_sn
  return(result)
}

#Creating a program that will loop through and add susceptibility

replicate_model <- function(total_weeks = c(), S_0, I_a, I_s, b_a, b_s, p0, c_a, c_s, T_pcr, T_rap, gamma, TFN_a, TFN_s, TISO_an, TISO_sn, TISO_ap, TISO_sp, FNpcr_a, FNpcr_s, FNrap_a, FNrap_s, num_weeks) {
  S_dot = s_dot(b_a, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn)
  I_a_dot = I_a_dot(p0, b_a, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn, T_pcr, T_rap, gamma)
  I_s_dot = I_s_dot(p0, b_a, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn, T_pcr, T_rap, gamma)
  TFN_a_dot = TFN_a_dot(T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TFN_a)
  TFN_s_dot = TFN_s_dot(T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TFN_s)
  TISO_ap_dot = TISO_ap_dot(c_a, T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TISO_ap)
  TISO_sp_dot = TISO_sp_dot(c_s, T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TISO_sp)
  TISO_an_dot = TISO_an_dot(c_a, T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TISO_an)
  TISO_sn_dot = TISO_sn_dot(c_s, T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TISO_sn)
  
  weekly_prediction <- c(S_0 + S_dot, I_a + I_a_dot, I_s + I_s_dot, TFN_a + TFN_a_dot, TFN_s + TFN_s_dot, TISO_ap + TISO_ap_dot, TISO_sp + TISO_sp_dot, TISO_an + TISO_an_dot, TISO_sn + TISO_sn_dot)
  total_weeks = append(total_weeks, weekly_prediction)

# may need to adjust the while function -- think about the logic here 
  
  if(num_weeks == 1) {
    return(total_weeks)
  } else {
    return(replicate_model(total_weeks, S_0 + S_dot, I_a + I_a_dot, I_s + I_s_dot, b_a, b_s, p0, c_a, c_s, T_pcr, T_rap, gamma, TFN_a + TFN_a_dot, TFN_s + TFN_s_dot, TISO_an + TISO_an_dot, TISO_sn + TISO_sn_dot, TISO_ap + TISO_ap_dot, TISO_sp + TISO_sp_dot, FNpcr_a, FNpcr_s, FNrap_a, FNrap_s, num_weeks - 1))
  }
}

result = replicate_model(total_weeks = c(), 0.895, 0.0025, 0.0025, 0.1, 0.2, 0.5, 0.7, 0.9, 1, 0, 0.1, 0, 0, 0, 0, 0, 0, 0.15, 0.15, 0.30, 0.30, 5); result

```

Now we run the simulation against various testing availabilities, beginning with the baseline test (no testing available).

```{r}
num_weeks = 15

T_pcr = 0.00
T_rap = 0.00

baseline <- replicate_model(total_weeks = c(), S_0, I_a, I_s, b_a, b_s, p0, c_a, c_s, T_pcr, T_rap, gamma, TFN_a, TFN_s, TISO_an, TISO_sn, TISO_ap, TISO_sp, FNpcr_a, FNpcr_s, FNrap_a, FNrap_s, num_weeks)
baseline <- t(matrix(baseline, ncol=num_weeks))
baseline
sums <- colSums(baseline); sums
0.04938909 + 0.04938909

T_pcr = 1.00
T_rap = 0.00

PCR_100 <- replicate_model(total_weeks = c(), S_0, I_a, I_s, b_a, b_s, p0, c_a, c_s, T_pcr, T_rap, gamma, TFN_a, TFN_s, TISO_an, TISO_sn, TISO_ap, TISO_sp, FNpcr_a, FNpcr_s, FNrap_a, FNrap_s, num_weeks)
PCR_100 <- t(matrix(PCR_100, ncol=num_weeks))

T_pcr = 0.50
T_rap = 0.00

PCR_50 <- replicate_model(total_weeks = c(), S_0, I_a, I_s, b_a, b_s, p0, c_a, c_s, T_pcr, T_rap, gamma, TFN_a, TFN_s, TISO_an, TISO_sn, TISO_ap, TISO_sp, FNpcr_a, FNpcr_s, FNrap_a, FNrap_s, num_weeks)
PCR_50 <- t(matrix(PCR_50, ncol=num_weeks))

T_pcr = 0.33
T_rap = 0.00

PCR_33 <- replicate_model(total_weeks = c(), S_0, I_a, I_s, b_a, b_s, p0, c_a, c_s, T_pcr, T_rap, gamma, TFN_a, TFN_s, TISO_an, TISO_sn, TISO_ap, TISO_sp, FNpcr_a, FNpcr_s, FNrap_a, FNrap_s, num_weeks)
PCR_33 <- t(matrix(PCR_33, ncol=num_weeks))

T_pcr = 0.00
T_rap = 1.00

RAP_100 <- replicate_model(total_weeks = c(), S_0, I_a, I_s, b_a, b_s, p0, c_a, c_s, T_pcr, T_rap, gamma, TFN_a, TFN_s, TISO_an, TISO_sn, TISO_ap, TISO_sp, FNpcr_a, FNpcr_s, FNrap_a, FNrap_s, num_weeks)
RAP_100 <- t(matrix(RAP_100, ncol=num_weeks))

T_pcr = 0.00
T_rap = 0.50

RAP_50 <- replicate_model(total_weeks = c(), S_0, I_a, I_s, b_a, b_s, p0, c_a, c_s, T_pcr, T_rap, gamma, TFN_a, TFN_s, TISO_an, TISO_sn, TISO_ap, TISO_sp, FNpcr_a, FNpcr_s, FNrap_a, FNrap_s, num_weeks)
RAP_50 <- t(matrix(RAP_50, ncol=num_weeks))

T_pcr = 0.00
T_rap = 0.33

RAP_33 <- replicate_model(total_weeks = c(), S_0, I_a, I_s, b_a, b_s, p0, c_a, c_s, T_pcr, T_rap, gamma, TFN_a, TFN_s, TISO_an, TISO_sn, TISO_ap, TISO_sp, FNpcr_a, FNpcr_s, FNrap_a, FNrap_s, num_weeks)
RAP_33 <- t(matrix(RAP_33, ncol=num_weeks))
```

The following attempts to use the deSolve package in order to solve the ODEs from the paper as instructed by April Yu. 

```{r}
library(deSolve)
library(scatterplot3d)
library(lubridate)

parameters <- c(b_a = 0.1, b_s = 0.2, p0 = 0.5, c_a = 0.7, c_s = 0.9, T_pcr = 1, T_rap = 0, gamma = 0.1, FNpcr_a = 0.15, FNpcr_s = 0.15, FNrap_a = 0.3, FNrap_s = 0.3)
state <- c(S_0 = 0.895, I_a = 0.0025, I_s = 0.0025, TFN_a = 0, TFN_s = 0, TISO_an = 0, TISO_sn = 0, TISO_ap = 0, TISO_sp = 0)
replicate_function2 <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    dS <- (-b_a * S_0 * (I_a + TFN_a + TISO_an)) - (b_s * S_0 * (I_s + TFN_s + TISO_sn))
    dI_a <- (1 - p0) * (b_a * S_0 * (I_a + TFN_a + TISO_an) + b_s * S_0 * (I_s + TFN_s + TISO_sn)) - T_pcr * I_a - T_rap * I_a - gamma * I_a
    dI_s <- p0 * (b_a * S_0 * (I_a + TFN_a + TISO_an) + b_s * S_0 * (I_s + TFN_s + TISO_sn)) - T_pcr * I_s - T_rap * I_s - gamma * I_s
    dTFN_a <- T_pcr * FNpcr_a * I_a + T_rap * FNrap_a * I_a - gamma * TFN_a
    dTFN_s <- T_pcr * FNpcr_s * I_s + T_rap * FNrap_s * I_s - gamma * TFN_s
    dTISO_ap<- c_a * (T_pcr * FNpcr_a * I_a + T_rap * FNrap_a * I_a) - gamma * TISO_ap
    dTISO_sp <- c_s * (T_pcr * FNpcr_s * I_s + T_rap * FNrap_s * I_s) - gamma * TISO_sp
    dTISO_an<- (1 - c_a) * (T_pcr * FNpcr_a * I_a + T_rap * FNrap_a * I_a) - gamma * TISO_an
    dTISO_sn <- (1 - c_s) * (T_pcr * FNpcr_s * I_s + T_rap * FNrap_s * I_s) - gamma * TISO_sn
    return(list(c(dS, dI_a, dI_s, dTFN_a, dTFN_s, dTISO_ap, dTISO_sp, dTISO_an, dTISO_sn)))
  })
}

sir_start_date <- "2021-01-18"
sir_end_date <- "2021-05-05"
#times <- 1:as.integer(ymd(sir_end_date) + 1 - ymd(sir_start_date))

times <- seq(0, 107, by=7)
out <- ode(y = state, times = times, func = replicate_function2, parms = parameters)
out
par(oma = c(0, 0, 3, 0))
plot(out, xlab = "time", ylab = "-")
sums4 <- colSums(out); sums4


```
