---
title: "Model Replication"
author: "Kyle Rohde"
date: "9/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The purpose of this document is to replicate "The Impact of Covid-19 Testing on College Campuses" research done by Zhuoting Yu, Akane B. Fujimoto, Pinar Keskinocak, and Julie Swann.

The model created considers various states to describe a student's status as it relates to Covid-19 infection. The states are as follows:

$S$: susceptible

$I_a$: infected and asymptomatic

$I_s$: infected and symptomatic

$TFN_{a(s)}$: infected and asymptomatic (symptomatic), has been tested with false negative results

$TISO^+_{a(s)}$: infected and asymptomatic (symptomatic), has been tested and currently in isolation

$TISO^-_{a(s)}$: infected and asymptomatic (symptomatic), has been tested and currently not in isolation

$RC$: recovered and confirmed, i.e., has been diagnosed previously

$RU$: recovered and unknown, i.e., has not been diagnosed previously

The follow equations describe the fraction of individuals in each state over time

$$
S = -\beta_aS(I_a + TFN_a + TISO_a^-)-\beta_sS(I_s+TFN_s+TISO_s^-)
$$

$$
I_a= (1-p_0) \times [\beta_aS(I_a+TFN_a+TISO_a^-)+\beta_sS(I_s+TFN_s+TISO_s^-)]-T_{pcr} \times I_a-T{rap} \times I_a -\gamma_aI_a
$$

$$
I_s = p_0 \times [\beta_aS(I_a+TFN_a+TISO_a^-)+\beta_sS(I_s+TFN_s+TISO_s^-)] − T_{pcr} × I_s − T_{rap} × I_s − γ_sI_s
$$

$$
TFN_a =T_{pcr} × FN_{pcr_a} × I_a + T_{rap} × FN_{rap_a} × I_a − γ_a × TFN_a
$$

$$
TFN_s =T_{pcr} × FN_{pcr_s} × I_s + T_{rap} × FN_{rap_s} × I_s − γ_s × TFN_s
$$

$$
TISO^+_a =C_a × [T_{pcr} × FN_{pcr_a} × I_a + T_{rap} × FN_{rap_a} × I_a] − γ_a × TISO^+_a
$$

$$
TISO^+_s =C_s × [T_{pcr} × FN_{pcr_s} × I_s + T_{rap} × FN_{rap_s} × I_s] − γ_s × TISO^+_s
$$

$$
TISO^−_a =(1 − C_a) × [T_{pcr} × FN_{pcr_a} × I_a + T_{rap} × FN_{rap_a} × I_a] − γ_a × TISO^−_a
$$

$$
TISO^−_s =(1 − C_s) × [T_{pcr} × FN_{pcr_s} × I_s + T_{rap} × FN_{rap_s} × I_s] − γ_s × TISO^−_s
$$

Below are the variable definitions

|          Parameter         |                                             Description                                             |           Value          |
|:--------------------------:|:---------------------------------------------------------------------------------------------------:|:------------------------:|
|     $\beta_a$,$\beta_s$    | Transmission rate due to contact between susceptible and asymptomatic/symptomatic infected subjects |          0.1,0.2         |
|          $\gamma$          |                              Rate of recovery of an infected individual                             |            0.1           |
|            $p_0$           |                            Proportion of infections that are symptomatic                            |            0.5           |
|         $C_a$,$C_s$        |        Isolation compliance rate for asymptomatic/symptomatic infected and tested individuals       |         0.7, 0.9         |
|  $FN_{pcr_a}$,$FN_{pcr_s}$ |               False negative rate of a PCR test for asymptomatic/symptomatic patients               |         15%, 15%         |
| $FN_{rap_a}$, $FN_{rap_s}$ |              False negative rate of a rapid test for asymptomatic/symptomatic patients              |         30%, 30%         |
| $FP_{pcr_a}$, $FP_{pcr_s}$ |               False positive rate of a PCR test for asymptomatic/symptomatic patients               |          5%, 5%          |
| $FP_{rap_a}$, $FP_{rap_s}$ |               False positive ate of a rapid test for asymptomatic/symptomatic patients              |         10%, 10%         |
|          $T_{pcr}$         |                         Percentage of students that have access to PCR tests                        | Defined in each scenario |
|  $T_{rap_a}$, $T_{rap_s}$  |                    Percentage of students that have access to rapid antigen tests                   | Defined in each scenario |

* Baseline: No testing
* $PCR_{100}$: PCR tests available to 100% of eligible student body per week
* $RA_{100}$: Rapid antigen tests available to 100% of the eligible student body per week
* $PCR_{50}$: PCR tests available to 50% of the eligible student body per week
* $RA_{50}$: Rapid antigen tests available to 50% of the eligible student body per week
* $PCR_{33}$: PCR tests available to 33% of the eligible student body per week
* $RA_{33}$: Rapid antigen tests available to 33% of the eligible student body per week

```{r}
# Constants
ba = 0.1        # transmission rate due to contact of asymptomatic patient
bs = 0.2        # transmission rate due to contact of symptomatic patient
gamma = 0.1     # rate of recovery for an infected individual
p0 = 0.5        # proportion of infections that are symptomatic
ca = 0.7        # isolation compliance asymptomatic infected + tested patients
cs = 0.9        # isolation compliance symptomatic infected + tested patients
FNpcr_a = 0.15  # false negative rate of a PCR test for asymptomatic patients
FNpcr_s = 0.15  # false negative rate of a PCR test for symptomatic patients
FNrap_a = 0.30  # false negative rate of a rapid test for asymptomatic patients
FNrap_s = 0.30  # false negative rate of a rapid test for symptomatic patients
FPpcr_a = 0.05  # false positive rate of a PCR test for asymptomatic patients
FPpcr_s = 0.05  # false positive rate of a PCR test for symptomatic patients
FPrap_a = 0.10  # false positive rate of a rapid test for asymptomatic patients
FPrap_s = 0.10  # false positive rate of a rapid test for symptomatic patients

# Initial values
S = 0.895       # initial susceptibility of the student body
I_s = 0.0025    # initial infected and symptomatic
I_a = 0.0025    # initial infected and asymptomatic
R_u = 0.10      # initial recovered unknown

n = 15000       # student population
```

This is a test.






