---
title: "Model Replication"
author: "Kyle Rohde, Caroline Devlin"
date: "9/10/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The purpose of this document is to replicate "The Impact of Covid-19 Testing on College Campuses" research done by Zhuoting Yu, Akane B. Fujimoto, Pinar Keskinocak, and Julie Swann.

The model created considers various states to describe a student's status as it relates to Covid-19 infection. The states are as follows:

$S$: susceptible

$I_a$: infected and asymptomatic

$I_s$: infected and symptomatic

$TFN_{a(s)}$: infected and asymptomatic (symptomatic), has been tested with false negative results

$TISO^+_{a(s)}$: infected and asymptomatic (symptomatic), has been tested and currently in isolation

$TISO^-_{a(s)}$: infected and asymptomatic (symptomatic), has been tested and currently not in isolation

$RC$: recovered and confirmed, i.e., has been diagnosed previously

$RU$: recovered and unknown, i.e., has not been diagnosed previously

The follow equations describe the fraction of individuals in each state over time

$$
S = -\beta_aS(I_a + TFN_a + TISO_a^-)-\beta_sS(I_s+TFN_s+TISO_s^-)
$$

$$
I_a= (1-p_0) \times [\beta_aS(I_a+TFN_a+TISO_a^-)+\beta_sS(I_s+TFN_s+TISO_s^-)]-T_{pcr} \times I_a-T{rap} \times I_a -\gamma_aI_a
$$

$$
I_s = p_0 \times [\beta_aS(I_a+TFN_a+TISO_a^-)+\beta_sS(I_s+TFN_s+TISO_s^-)] − T_{pcr} × I_s − T_{rap} × I_s − γ_sI_s
$$

$$
TFN_a =T_{pcr} × FN_{pcr_a} × I_a + T_{rap} × FN_{rap_a} × I_a − γ_a × TFN_a
$$

$$
TFN_s =T_{pcr} × FN_{pcr_s} × I_s + T_{rap} × FN_{rap_s} × I_s − γ_s × TFN_s
$$

$$
TISO^+_a =C_a × [T_{pcr} × FN_{pcr_a} × I_a + T_{rap} × FN_{rap_a} × I_a] − γ_a × TISO^+_a
$$

$$
TISO^+_s =C_s × [T_{pcr} × FN_{pcr_s} × I_s + T_{rap} × FN_{rap_s} × I_s] − γ_s × TISO^+_s
$$

$$
TISO^−_a =(1 − C_a) × [T_{pcr} × FN_{pcr_a} × I_a + T_{rap} × FN_{rap_a} × I_a] − γ_a × TISO^−_a
$$

$$
TISO^−_s =(1 − C_s) × [T_{pcr} × FN_{pcr_s} × I_s + T_{rap} × FN_{rap_s} × I_s] − γ_s × TISO^−_s
$$

Below are the variable definitions

|          Parameter         |                                             Description                                             |           Value          |
|:--------------------------:|:---------------------------------------------------------------------------------------------------:|:------------------------:|
|     $\beta_a$,$\beta_s$    | Transmission rate due to contact between susceptible and asymptomatic/symptomatic infected subjects |          0.1,0.2         |
|          $\gamma$          |                              Rate of recovery of an infected individual                             |            0.1           |
|            $p_0$           |                            Proportion of infections that are symptomatic                            |            0.5           |
|         $C_a$,$C_s$        |        Isolation compliance rate for asymptomatic/symptomatic infected and tested individuals       |         0.7, 0.9         |
|  $FN_{pcr_a}$,$FN_{pcr_s}$ |               False negative rate of a PCR test for asymptomatic/symptomatic patients               |         15%, 15%         |
| $FN_{rap_a}$, $FN_{rap_s}$ |              False negative rate of a rapid test for asymptomatic/symptomatic patients              |         30%, 30%         |
| $FP_{pcr_a}$, $FP_{pcr_s}$ |               False positive rate of a PCR test for asymptomatic/symptomatic patients               |          5%, 5%          |
| $FP_{rap_a}$, $FP_{rap_s}$ |               False positive ate of a rapid test for asymptomatic/symptomatic patients              |         10%, 10%         |
|          $T_{pcr}$         |                         Percentage of students that have access to PCR tests                        | Defined in each scenario |
|  $T_{rap_a}$, $T_{rap_s}$  |                    Percentage of students that have access to rapid antigen tests                   | Defined in each scenario

```{r}
library("readxl")

# load dataset & remove unnecessary columns
data <- read.csv("county_cases.csv", header = TRUE)
attach(data)

county_data <- subset(data, select = -c(county_id, 
                                 State.FIPS.code,
                                 County.FIPS.code,
                                 hospitalization,
                                 deaths,
                                 case.rate,
                                 death.rate,
                                 X14.day.case.rate,
                                 antigen_cases,
                                 probable_deaths))
head(county_data)
```

```{r}
library(hash)

counties <- c(
              "Appling", "Atkinson", "Bacon", "Baker", "Baldwin", "Banks", "Barrow", "Bartow",
              "Ben Hill", "Berrien", "Bibb", "Bleckley", "Brantley", "Brooks", "Bryan", "Bulloch",
              "Burke", "Butts", "Calhoun", "Camden", "Candler", "Carroll", "Catoosa", "Charlton",
              "Chatham", "Chattahoochee", "Chattooga", "Cherokee", "Clarke", "Clay", "Clayton",
              "Clinch", "Cobb", "Coffee", "Colquitt", "Columbia", "Cook", "Coweta", "Crawford", "Crisp",
              "Dade", "Dawson", "DeKalb", "Decatur", "Dodge", "Dooly", "Dougherty", "Douglas", "Early",
              "Echols", "Effingham", "Elbert", "Emanuel", "Evans", "Fannin", "Fayette", "Floyd",
              "Forsyth", "Franklin", "Fulton", "Gilmer", "Glascock", "Glynn", "Gordon", "Grady",
              "Greene", "Gwinnett", "Habersham", "Hall", "Hancock", "Haralson", "Harris", "Hart",
              "Heard", "Henry", "Houston", "Irwin", "Jackson", "Jasper", "Jeff Davis", "Jefferson",
              "Jenkins", "Johnson", "Jones", "Lamar", "Lanier", "Laurens", "Lee", "Liberty", "Lincoln",
              "Long", "Lowndes", "Lumpkin", "Macon", "Madison", "Marion", "McDuffie", "McIntosh", 
              "Meriwether", "Miller", "Mitchell", "Monroe", "Montgomery", "Morgan", "Murray", 
              "Muscogee", "Newton", "Oconee", "Oglethorpe", "Paulding", "Peach", "Pickens", "Pierce", 
              "Pike", "Polk", "Pulaski", "Putnam", "Quitman", "Rabun", "Randolph", "Richmond", 
              "Rockdale", "Schley", "Screven", "Seminole", "Spalding", "Stephens", "Stewart", "Sumter",
              "Talbot", "Taliaferro", "Tattnall", "Taylor", "Telfair", "Terrell", "Thomas", "Tift", 
              "Toombs", "Towns", "Treutlen", "Troup", "Turner", "Twiggs", "Union", "Upson", "Walker",
              "Walton", "Ware", "Warren", "Washington", "Wayne", "Webster", "Wheeler", "White", 
              "Whitfield", "Wilcox", "Wilkes", "Wilkinson", "Worth"
             )

generate_default_school <- function() {
  school <- hash()
  school["county"]           <- ""
  school["days"]             <- 107
  school["masking"]          <- 0
  school["student_pop"]      <- 0
  school["faculty_pop"]      <- 0
  school["student_vax"]      <- 0
  school["faculty_vax"]      <- 0
  school["testing"]          <- 0 # only considering rapid testing
  return(school)
}

```

The following calculates the initial states for each county using publicly available, county-level data from the DPH website and adds them to a data frame with the following column-headers for each county: County Name (county_name), Total Number of Cases per County (cases), Total Population per County (population), Number of New Cases in the Past 14 Days per County (X14.day.cases), Initial Susceptibility (S_0), Initial Infected Symptomatic (I_s), Initial Infected Asymptomatic (I_a), Initial Recovered Unknown (R_u), and the Total Sum of Initial States (total).

```{r}
initial_susceptibility <- function(county_data) {
  result <- (county_data$population - county_data$cases)/county_data$population
  return(result)
}

initial_infected_symptomatic <- function(county_data) {
  result <- ((county_data$X14.day.cases)/county_data$population)*0.5
  return(result)
}

initial_infected_asymptomatic <- function(county_data) {
  result <- ((county_data$X14.day.cases)/county_data$population)*0.5
  return(result)
}

initial_recovered_unknown <- function(county_data) {
  result <- (county_data$cases - county_data$X14.day.cases)/county_data$population
  return(result)
}

county_data$S_0 <- initial_susceptibility(county_data)
county_data$I_s <- initial_infected_symptomatic(county_data)
county_data$I_a <- initial_infected_asymptomatic(county_data)
county_data$R_u <- initial_recovered_unknown(county_data)

county_data$total <- county_data$S_0 + county_data$I_s + county_data$I_a + county_data$R_u

# We should exclude Non-GA Resident and Unknown


```

Below is the R code for defining the initial values and constants of the system.

```{r}
# Initial states
b_a = 0.1       # transmission rate due to contact of asymptomatic patient
b_s = 0.2       # transmission rate due to contact of symptomatic patient
gamma = 0.1     # rate of recovery for an infected individual
p0 = 0.5        # proportion of infections that are symptomatic
c_a = 0.7       # isolation compliance asymptomatic infected + tested patients
c_s = 0.9       # isolation compliance symptomatic infected + tested patients
FNpcr_a = 0.15  # false negative rate of a PCR test for asymptomatic patients
FNpcr_s = 0.15  # false negative rate of a PCR test for symptomatic patients
FNrap_a = 0.30  # false negative rate of a rapid test for asymptomatic patients
FNrap_s = 0.30  # false negative rate of a rapid test for symptomatic patients
FPpcr_a = 0.05  # false positive rate of a PCR test for asymptomatic patients
FPpcr_s = 0.05  # false positive rate of a PCR test for symptomatic patients
FPrap_a = 0.10  # false positive rate of a rapid test for asymptomatic patients
FPrap_s = 0.10  # false positive rate of a rapid test for symptomatic patients

# Initial values
S_0 = initial_susceptibility(county_data)     # initial susceptibility of the student body
I_s = initial_infected_symptomatic(county_data)    # initial infected and symptomatic
I_a = initial_infected_asymptomatic(county_data)    # initial infected and asymptomatic
R_u = initial_recovered_unknown(county_data)      # initial recovered unknown
R_c = 0.0       # initial recovered confirmed

# Initial test values (no tests administered)
TFN_a = 0       # initial population with a false negative and asymptomatic
TFN_s = 0       # initial population with a false negative and symptomatic 
TISO_an = 0     # initial population tested with infection, not in isolation, asymptomatic 
TISO_ap = 0     # initial population tested with infection, in isolation, asymptomatic 
TISO_sn = 0     # initial population tested with infection, not in isolation, symptomatic
TISO_sp = 0     # initial population tested with infection, in isolation, symptomatic

n = 15000       # student population

# Days in simulation
days <- 107

# Sequence for semester
semester <- seq(1, days, by=1)

# Initial state of the system
initial_state <- c(S_0 = S_0, I_a = I_a, I_s = I_s, TFN_a = TFN_a, TFN_s = TFN_s, TISO_an = TISO_an, TISO_sn = TISO_sn, TISO_ap = TISO_ap, TISO_sp = TISO_sp, RC = R_c, RU = R_u)
```

The following code defines the differential equations for the change in the states over time. The functions allow the user to input current values for the states and returns the derivative at that time step. 

```{r}
S_dot <- function(b_a, MC, VR_s, VR_f, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn) {
  result <- (-b_a * (1 - MC) * (1 - (Vr_s + VR_f)) * S_0 * (I_a + TFN_a + TISO_an)) - (b_s * (1 - MC) * (1 - VR) * S_0 * (I_s + TFN_s + TISO_sn))
  return(result)
}

I_a_dot <- function(p0, b_a, MC, VR_s, VR_f, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn, T_pcr, T_rap, gamma) {
  result <- (1 - p0) * ((b_a * (1 - MC) * (1 - (Vr_s + VR_f)) * S_0 * (I_a + TFN_a + TISO_an)) + (b_s * (1 - MC) * (1 - VR) * S_0 * (I_s + TFN_s + TISO_sn))) - (T_pcr * I_a) - (T_rap * I_a) - (gamma * I_a)
  return(result)
}

I_s_dot <- function(p0, b_a, MC, VR_s, VR_f, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn, T_pcr, T_rap, gamma) {
  result <- p0 * ((b_a * (1 - MC) * (1 - (Vr_s + VR_f)) * S_0 * (I_a + TFN_a + TISO_an)) + (b_s * (1 - MC) * (1 - VR) * S_0 * (I_s + TFN_s + TISO_sn))) - (T_pcr * I_s) - (T_rap * I_s) - (gamma * I_s)
  return(result)
}

TFN_a_dot <- function(T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TFN_a) {
  result <- (T_pcr * FNpcr_a * I_a) + (T_rap * FNrap_a * I_a) - (gamma * TFN_a)
  return(result)
}

TFN_s_dot <- function(T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TFN_s) {
  result <- (T_pcr * FNpcr_s * I_s) + (T_rap * FNrap_s * I_s) - (gamma * TFN_s)
  return(result)
}

TISO_ap_dot <- function(c_a, T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TISO_ap) {
  result <- c_a * ((T_pcr * (1 -FNpcr_a) * I_a) + (T_rap * (1 - FNrap_a) * I_a)) - (gamma * TISO_ap)
  return(result)
}

TISO_sp_dot <- function(c_s, T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TISO_sp) {
  result <- c_s * ((T_pcr * (1 - FNpcr_s) * I_s) + (T_rap * (1 - FNrap_s) * I_s)) - (gamma * TISO_sp)
  return(result)
}

TISO_an_dot <- function(c_a, T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TISO_an) {
  result <- (1 - c_a) * ((T_pcr * (1 - FNpcr_a) * I_a) + (T_rap * (1 - FNrap_a) * I_a)) - (gamma * TISO_an)
  return(result)
}

TISO_sn_dot <- function(c_s, T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TISO_sn) {
  result <- (1 - c_s) * ((T_pcr * (1 - FNpcr_s) * I_s) + (T_rap * (1 - FNrap_s) * I_s)) - (gamma * TISO_sn)
  return(result)
}

RC_dot <- function(gamma, TISO_ap, TISO_sp, TISO_an, TISO_sn) {
  result <- gamma * (TISO_ap + TISO_sp + TISO_an + TISO_sn)
  return(result)
}

RU_dot <- function(gamma, I_a, I_s, TFN_a, TFN_s) {
  result <- gamma * (I_a + I_s + TFN_a + TFN_s)
  return(result)
}
```

Now that the system's equations have been defined, we define functions to observe the changes in the system. These functions include all of the state derivative equations as well as a function that loops through the days, solving the differential of all equations for each day of the semester. The function `run_simulation` takes 

* `state`: The initial state of the system,
* `testing`: The parameters for a day in which testing occurs
* `no_testing`: The parameters for a day in which no testing occurs
* `semester`: A time sequence representing the simulation semester

and returns a list containing the values of each state for each day. 

```{r, results='hide'}
library(deSolve)
library(scatterplot3d)
library(lubridate)
```

```{r}
# Define generic differential update function
differentials <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    dS <- S_dot(b_a, MC, VR_s, VR_f, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn)
    dI_a <- I_a_dot(p0, b_a, MC, VR_s, VR_f, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn, T_pcr, T_rap, gamma)
    dI_s <- I_s_dot(p0, b_a, MC, VR_s, VR_f, S_0, I_a, TFN_a, TISO_an, b_s, I_s, TFN_s, TISO_sn, T_pcr, T_rap, gamma)
    dTFN_a <- TFN_a_dot(T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TFN_a)
    dTFN_s <- TFN_s_dot(T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TFN_s)
    dTISO_ap <- TISO_ap_dot(c_a, T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TISO_ap)
    dTISO_sp <- TISO_sp_dot(c_s, T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TISO_sp)
    dTISO_an <- TISO_an_dot(c_a, T_pcr, FNpcr_a, I_a, T_rap, FNrap_a, gamma, TISO_an)
    dTISO_sn <- TISO_sn_dot(c_s, T_pcr, FNpcr_s, I_s, T_rap, FNrap_s, gamma, TISO_sn)
    dRC <- RC_dot(gamma, TISO_ap, TISO_sp, TISO_an, TISO_sn)
    dRU <- RU_dot(gamma, I_a, I_s, TFN_a, TFN_s)
    return(list(c(dS, dI_a, dI_s, dTFN_a, dTFN_s, dTISO_an, dTISO_sn, dTISO_ap, dTISO_sp, dRC, dRU)))
  })
}

run_simulation <- function(state, testing, no_testing, semester) {
  states <- list()
  for (t in semester) {
    day <- seq(t - 1, t, by=1)
    if (t %% 7 == 0) {
      parameters <- testing
    } else {
      parameters <- no_testing
    }
    state <- ode(y = state, times = day, func = differentials, parms = parameters)[2,c(2:12)]
    states[[length(states) + 1]] <- state
  }
  return(states)
}
```


```{r}
# Input county and run simulation
# Kyle may be working on this 

```

```{r}
# Helper code to generate scenarios

# Rapid Testing: 0, 0.25, 0.5, 0.75, 1
# Student Vaccination: 0, 0.25, 0.5, 0.75, 1
# Faculty Vaccination: 0, 0.25, 0.5, 0.75, 1
# Masking Compliance (Translated to Masking Efficacy): 0, 0.25, 0.5, 0.75, 1

rapid_testing <- seq(0, 1, by=0.25)
student_vaccination <- seq(0, 1, by=0.25)
faculty_vaccination <- seq(0, 1, by=0.25)
masking_compliance <- seq(0, 1, by=0.25)

i = 1

for (s in student_vaccination) {
  for (f in faculty_vaccination) {
    for (m in masking_compliance) {
      for (r in rapid_testing) {
        result <- sprintf("testing%d <- c(b_a = b_a, b_s = b_s, p0 = p0, c_a = c_a, c_s = c_s, MC = %f, VR_s = %f, VR_f = %f, T_pcr = 0, T_rap = %f, gamma = gamma, FNpcr_a = FNpcr_a, FNpcr_s = FNpcr_s, FNrap_a = FNrap_a, FNrap_s = FNrap_s) \n no_testing%d <- c(b_a = b_a, b_s = b_s, p0 = p0, c_a = c_a, c_s = c_s, MC = %f, VR_s = %f, VR_f = %f, T_pcr = 0, T_rap = 0, gamma = gamma, FNpcr_a = FNpcr_a, FNpcr_s = FNpcr_s, FNrap_a = FNrap_a, FNrap_s = FNrap_s) \n scenario%d <- run_simulation(initial_state, testing%d, no_testing%d, semester) \n", i, m, s, f, r, i, m, s, f, i, i, i)
        cat(result)
        i = i + 1
      }
    }
  }
}

```


